---
title: "16-17 DPH101 resit C. Comparing Proportions"
author: "dizhen"
date: "2020.1.5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## C. Comparing Proportions
### Step1. Data

```{r}
preference <- matrix(c(19,14,9,45,5,10,6,39),ncol=4,byrow=TRUE)
colnames(preference) <- c("Day1","Day2","Day3","Day4")
rownames(preference) <- c("Black Oil","Striped")
mytable <- as.table(preference)
mytable
```

### Step2. Assumptions
1. The sample data are randomly selected.
2. The sample data are represented in two-way table
3. No more than 20% of the cells in the contingency table have expected frequencies < 5.(The lowest expected frequency is 8.878) The requirements are satisfied.

```{r}
chisq.test(mytable,corr = F)$expected
```
### Step3. Hypothesis

H0: The preference for sunflower seeds and different days are independent 

H1: The preference for sunflower seeds and different days are dependent

### Step4. Test statistics
### Step5. Distribution of the Test Statistic

We use $\chi^2 ~~distribution$


### Step6. Decision rule

### Step7. Calculation of the test statistic

Chi-square test

```{r}
chisq.test(mytable, corr = F)
```

### Step8. Statistical decision

### Step9. Conclusion

As p-value (=0.167) is large (larger than 0.05), we fail to reject the null hypothesis and conclude that the preference for sunflower seeds and different days are independent.  

### Step10. p-value 

1. p-value is 0.167, there is a nearly 16.7% chance of observing a dependence from randomly sampling as large or larger than observed. 

2. If you repeated the study many times, nearly 16.7% of the time you would observe the same or larger dependence.
