---
title: "18/19 DPH101 final D. Comparing Proportions"
author: "dizhen"
date: "2019.12.27"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## D. Comparing Proportions
### Step1. Data
```{r}
percent <- c(15.2, 30.4,24.2,30.2, 5.8,23.6,40.8,29.8,0.2,0.4,1.2,98.2,59.8,26.8,11.8,1.6)
n_percent <- percent * 5
level <- factor(c(rep(c(0:3), times = 4)),
              labels = c("4", "3", "2", "1"),
              ordered = F)
location <- factor(c(rep(0, times = 4), 
                 rep(1, times = 4), 
                 rep(2, times = 4),
                 rep(3, times = 4)),
               labels = c("Qiongliu", "Gaomi", "Jiange","Zichuan"))
mydata <- data.frame(n_percent,level, location )
mytable <- xtabs(n_percent~location+level, mydata)
mytable

# Alternative
# percent <- matrix(c(15.2, 30.4,24.2,30.2, 5.8,23.6,40.8,29.8,0.2,0.4,1.2,98.2,59.8,26.8,11.8,1.6),ncol = 4, byrow = TRUE)
# colnames(percent) <- c("4", "3", "2", "1")
# rownames(percent) <- c("Qiongliu", "Gaomi", "Jiange","Zichuan")
# mytable <- as.table(percent)
# mytable

```
### Step2. Assumptions
1. The sample data are randomly selected.
2. The sample data are represented in two-way table
3. No more than 20% of the cells  in the contingency table have expected frequencies < 5.(The lowest expected frequency is 6.122)
The requirements are satisfied.

```{r}
chisq.test(mytable,corr = F)$expected
```


### Step3. Hypothesis

H0: The location and selenium levels are independent

H1: The location and selenium levels are dependent

### Step4. Test statistics
### Step5. Distribution of the Test Statistic

We use $\chi^2 ~~distribution$

### Step6. Decision rule

### Step7. Calculation of the test statistic

Chi-square test

```{r}
chisq.test(mytable, corr = F)
```

### Step8. Statistical decision

### Step9. Conclusion

As p-value (< 2.2e-16) is very small, we reject the null hypothesis and conclude that location and selenium levels are dependent.

### Step10. p-value

1. p-value is less than 2.2e-16, there is a nearly 0 chance of observing a dependence from randomly sampling as large or larger than observed.

2. If you repeated the study many times, nearly all the time you would observe smaller dependence and nearly 0 of the time you would observe larger dependence.



