---
title: "18/19 DPH101 final C.Compare More Than Two Means"
author: "dizhen"
date: "2019.12.27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## C.Compare More Than Two Means
### Step1. Data
```{r}
LEVELS <- c(0.637,0.354,0.287,0.533,0.228,0.632,0.303,0.194,0.293,0.276,0.226,0.395,0.994,0.89,0.809,0.321,0.712,0.949,0.933,0.28,0.336,0.678,0.714,0.757)
TEMPERATURE <- factor(c(rep(0, times = 6), rep(1, times=6),
                  rep(0, times = 6), rep(1, times = 6)),
                labels = c("Unheated", "Heated"))

SOURCES <- factor(c(rep(0, times = 12), rep(1, times=12)),
                labels = c("Juice", "Chromoplast"))
mydata <- data.frame(LEVELS,TEMPERATURE,SOURCES)
str(mydata)
head(mydata)
```

```{r}
boxplot(LEVELS ~ TEMPERATURE * SOURCES,
        data = mydata)
```
Conclusion. There is no evidence of non-normality (boxplots not grossly asymmetrical), variances are a little unequal (although perhaps not grossly unequal - one of the boxplots is not more than three times smaller than the other). 

### Step2. Assumptions
1. For each group, the sample values appear to be from a population with a distribution that is approximately normal.
2. The variance of the groups as indicated in the graph differ considerably, but the test is robust against departures from equal variances.
3. The samples are simple random samples of subjects
4. The samples are independent of each other.
5. The sample values are categorized in two ways (temperature and source)
6. All of the groups have the same number(6) of sample values.
The requirements are satisfied. We perform two-way analysis of variance.

### Step3. Hypothesis
Because there seems to be an interaction effect, we perform factorial design.
$$H_0: There~~is~~no~~joint~~effect~~of~~source~~and~~temperature$$
$$H_A: There~~is~~an~~joint~~effect~~of~~source~~and~~temperature$$


### Step4. Test Statistic
F = MS(interaction)/MS(error)

### Step5. Distribution of the Test Statistic
We use F distribution

### Step6. Decision Rule
### Step7. Calculation of test statistic

```{r}
LEVELS.AOV <- aov(LEVELS ~ TEMPERATURE * SOURCES,
                    data = mydata)
summary(LEVELS.AOV)
```

### Step8. Statistical decision
### Step9. Conclusion
1. Interaction effect
We fail to reject the null hypothesis of no effect of interaction, as the p value (0.994) is large (more than 0.05). We conclude that there appears no effect of interaction of sources and temperature on the serum concentration of beta-carotene in ferrets.


### Step3. Hypothesis
$$H_0: There~~is~~no~~effect~~of~~a~~source$$
$$H_A: There~~is~~an~~effect~~of~~a~~source$$
$$H_0: There~~is~~no~~effect~~ of~~ temperature$$
$$H_A: There~~ is~~ an ~~effect ~~of ~~a ~~temperature$$



### Step4. Test Statistic
F = MS(source)/MS(error)
F = MS(temperature)/MS(error)

### Step5. Distribution of the Test Statistic
We use F distribution

### Step6. Decision Rule
### Step7. Calculation of test statistic

```{r}
LEVELS.AOV <- aov(LEVELS ~ TEMPERATURE * SOURCES,
                    data = mydata)
summary(LEVELS.AOV)
```

### Step8. Statistical decision
### Step9. Conclusion
2. Row/column effect
Only the null hypothesis of no effect of sources can be rejected, as the p-value (0.0006) of it is less than 0.05. We conclude that there is an effect of sources on the serum concentration of beta-carotene in ferrets.
We fail to reject the null hypothesis of no effect of temperature, as the p-value (0.061) is greater than 0.05. We conclude that there appears no effect of temperature on the serum concentration of beta-carotene in ferrets.

### Step10. P-value
1. There is a 0.06% chance of observing an effect of sources from randomly sampling subject as large or larger than observed.
2. If you repeated the study many times, 0.06% of the time you would observe larger effect.

### A single model
Write down a single model to assess the effect of temperature and source on serum concentration of beta-carotene, allowing for the posibility that differences between mean concentrations of different sources vary by temperature.

$$x_{ijk} = \mu + \alpha_i + \beta_j + (\alpha \beta)_{ij} + \varepsilon_{ijk}$$

where i = 1,2, j = 1,2, k = 1,2,3,4,5,6

$x_{ijk}$ is a typical observation

$\mu$ is an unknown constant

$\alpha_i$ is the effect due to sources

$\beta_i$ is the effect due to temperatures

$(\alpha \beta)_{ij}$ is the effect due to interaction of temperatures and sources

$\varepsilon_{ijk}$ is the error term representing all other sources of variation

### A statistical interaction
```{r}
Unh_Ju <- c(0.637,0.354,0.287,0.533,0.228,0.632)
h_Ju <- c(0.303,0.194,0.293,0.276,0.226,0.395)
Unh_Chr <- c(0.994,0.89,0.809,0.321,0.712,0.949)
h_Chr <- c(0.933,0.28,0.336,0.678,0.714,0.757)


beta <- matrix(c(mean(Unh_Ju), mean(h_Ju), mean(Unh_Chr), mean(h_Chr)) ,ncol = 2, byrow = TRUE)
colnames(beta) <- c("Unheated","Heated")
rownames(beta) <- c("Juice", "Chromoplast")
mytable <- as.table(beta)
mytable


library(dplyr)
newdata <- mydata %>% group_by(SOURCES,TEMPERATURE) %>% summarize_all("mean")
newdata

```
```{r}
op <- par(mar = c(4, 4, 0.1, 0.1))
plot(LEVELS ~ as.factor(TEMPERATURE), 
     data = newdata, 
     type = "n", 
     axes = F, 
     xlab = "", 
     ylab = "")
with(subset(newdata, SOURCES == "Juice"), 
     points(LEVELS ~ as.factor(TEMPERATURE), 
            pch = 21, 
            type = "o", 
            lwd = 1))
with(subset(newdata, SOURCES == "Chromoplast"), 
     points(LEVELS ~ as.factor(TEMPERATURE), 
            pch = 16, 
            type = "o", 
            lwd = 1))

axis(1, cex.axis = 0.8)
mtext(text = "Temperature", 
      side = 1, 
      line = 2)
axis(2, cex.axis = 0.8, las = 1)
mtext(text = "Levels", 
      side = 2, 
      line = 2)
legend("topleft", 
       leg = c("Juice", "Chromoplast"), 
       lty = 0, 
       pch = c(21, 16), 
       bty = "n", 
       cex = 0.7)
box(bty = "l")
```
An interaction effect means that the effect of one factor depends on the other factor and it's shown by the lines in our interaction graph running parallel.
In this case, the effect for serum beta-carotene levels have no interact with sources. That is,
carrot juice affects levels similarly with carrot chromoplast.
