---
title: "Week 10"
author: "Elmer V Villanueva"
date: "April 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Management


## Install and load the packages `car, plyr`
```{r}
if (!require("car")) install.packages("car", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("car")

if (!require("MuMIn")) install.packages("MuMIn", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("MuMIn")
```

## Read the GLOW500 data from the `GLOW500.csv` file, importing only the following variables: SUB_ID, AGE, BMI, RATERISK, SMOKE
```{r}
GLOW_4 <- read.csv("GLOW500.csv")[ ,c('SUB_ID', 'AGE', 'BMI', 'PRIORFRAC', 'FRACTURE', 'FRACSCORE', 'RATERISK', 'SMOKE', 'PREMENO')]
str(GLOW_4)
head(GLOW_4)
```

###########################################################
###########################################################

# Factors as independent variables.
## Regress BMI on AGE
```{r}
BMI.LM1 <- lm(BMI ~ AGE, data = GLOW_4)
summary(BMI.LM1)
```
Interpret the intercept of AGE.

## Regress BMI on SMOKE
```{r}
BMI.LM2 <- lm(BMI ~ SMOKE, data = GLOW_4)
summary(BMI.LM2)
```
Interpret the intercept of SMOKE.

## Transform SMOKE to a factor variable
```{r}
GLOW_4$SMOKE.F <- factor(GLOW_4$SMOKE, labels = c("No", "Yes"))
summary(GLOW_4$SMOKE)
summary(GLOW_4$SMOKE.F)
```

## Regress BMI on SMOKE
```{r}
BMI.LM3 <- lm(BMI ~ SMOKE.F, data = GLOW_4)
summary(BMI.LM3)
```
Interpret the coefficient of SMOKE.F.

### Compare the results of a linear regression of BMI on SMOKE and a t-test of BMI by SMOKE.
```{r}
# t-test
t.test(BMI ~ SMOKE.F, data = GLOW_4, var.equal = TRUE)
# Regression
summary(BMI.LM3)
```

## Regress BMI on RATERISK.
```{r}
GLOW_4$RATERISK.F <- factor(GLOW_4$RATERISK, labels = c("Less", "Same", "More"))
BMI.LM4 <- lm(BMI ~ RATERISK.F, data = GLOW_4)
summary(BMI.LM4)
```
Interpret the coefficient of RATERISK.

### Compare the results of a regression of BMI on RATERISK and a one-way analysis of variance of BMI by RATERISK.
```{r}
### One-way ANOVA
oneway.test(BMI ~ RATERISK.F, data = GLOW_4, var.equal = TRUE)
### Regression
summary(BMI.LM4)
```

## Let's vary the reference value of the factor variable.
```{r}
GLOW_4$RATERISK.F <- relevel(GLOW_4$RATERISK.F, ref = 2)
BMI.LM5 <- lm(BMI ~ RATERISK.F, data = GLOW_4)
summary(BMI.LM5)
```
Interpret the coefficient of RATERISK.

## Compare the coefficients
```{r}
coefficients(BMI.LM4)
coefficients(BMI.LM5)
```

###########################################################
###########################################################

# Choosing the 'best' model.

## Read the GLOW500 data from the `GLOW500.csv` file, importing only the following variables: SUB_ID, AGE, BMI, PRIORFRAC, FRACSCORE, RATERISK, SMOKE, PREMENO
```{r}
GLOW_7 <- read.csv("GLOW500.csv")[ ,c('SUB_ID', 'AGE', 'BMI', 'PRIORFRAC', 'FRACSCORE', 'RATERISK', 'SMOKE', 'PREMENO')]
str(GLOW_7)
head(GLOW_7)
```

## Transform appropriate integer variables into factor variables.
```{r}
GLOW_7$PRIORFRAC <- factor(GLOW_7$PRIORFRAC, labels = c("No", "Yes"))
GLOW_7$RATERISK <- factor(GLOW_7$RATERISK, labels = c("Less", "Same", "More"))
GLOW_7$SMOKE <- factor(GLOW_7$SMOKE, labels = c("No", "Yes"))
GLOW_7$PREMENO <- factor(GLOW_7$PREMENO, labels = c("No", "Yes"))
summary(GLOW_7)
```

## Regress BMI on AGE, FRACSCORE, RATERISK, PRIORFRAC, SMOKE, PREMENO
```{r}
BMI.FULL <- lm(BMI ~ AGE + FRACSCORE + RATERISK + 
                   PRIORFRAC + SMOKE + PREMENO,
               data = GLOW_7)
summary(BMI.FULL)
```

## Choosing the 'best' model.
```{r}
summary(lm(BMI ~ AGE + FRACSCORE + RATERISK + PRIORFRAC + SMOKE + PREMENO,
             data = GLOW_7))
```


### WARNING!! Automating the process. WARNING!!
```{r}
options(na.action = "na.fail")
dredge(BMI.FULL, rank = "AIC")
```





