---
title: "Week 08"
author: "Elmer V Villanueva"
date: "April 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Management


## Install and load the package `car`
```{r}
if (!require("car")) install.packages("car", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("car")
```

## Read the GLOW500 data from the `GLOW500.csv` file, importing only the following variables: AGE, BMI, SMOKE, RATERISK 
```{r}
GLOW_4 <- read.csv("GLOW500.csv")[ ,c('AGE', 'BMI', 'SMOKE', 'RATERISK')]
str(GLOW_4)
head(GLOW_4)
```

```{r}
GLOW_4$SMOKE
```

## Transform SMOKE and RATERISK into factor variables. See GLOW Description for labels.
```{r}
GLOW_4$SMOKE.F <- factor(GLOW_4$SMOKE, labels = c("No", "Yes"))
GLOW_4$SMOKE <- NULL

GLOW_4$RATERISK.F <- factor(GLOW_4$RATERISK, labels = c("Less", "Same", "Greater"))
GLOW_4$RATERISK <- NULL

summary(GLOW_4)
```

# Produce a scatterplot of BMI on AGE
```{r}
scatterplot(BMI ~ AGE, data = GLOW_4)
```

*Conclusion.* No strong evidence of non-normality, non-linearity and non-heteroskedasticity.

# Fit a simple linear regression of BMI on AGE.
## Specify the linear model
```{r}
GLOW_4.LM1 <- lm(BMI ~ AGE, data = GLOW_4)
```

## Examine the parameter estimates
```{r}
summary(GLOW_4.LM1)
```

*Conclusion.* The model is highly statistically significant (p<0.001), but explains only about 5% of the total variability in BMI. For every increase in age, BMI decreases by 0.15 units.

## Produce confidence intervals for the parameter estimates
```{r}
confint(GLOW_4.LM1)
```

*Conclusion.* The model is highly statistically significant (p<0.001), but explains only about 5% of the total variability in BMI. For every increase in age, BMI decreases by 0.15 units **(95% CI -0.20, -0.09)**.

## Test the assumptions of the model
### Produce the predicted values
```{r}
GLOW_4$PRED <- predict(GLOW_4.LM1)
```

### Produce residuals
```{r}
GLOW_4$RESID <- residuals(GLOW_4.LM1)
```

### Test assumption 1: linearity of effect
```{r}
scatterplot(BMI ~ AGE, data = GLOW_4)
```

*Conclusion.* Strong evidence of linear effect.

### Test assumption 2: independence
This is not tested statistically.

### Test assumption 3: normality of residuals
```{r}
hist(GLOW_4$RESID, main = "Histogram of Residuals")
plot(density(GLOW_4$RESID), main = "Density of Residuals")
qqnorm(GLOW_4$RESID, main = "Quantile-Quantile Plot of Residuals")
shapiro.test(GLOW_4$RESID)
```

*Conclusion.* Residuals appear normally distributed, despite the statistically significant Shapiro-Wilk test.

### Test assumption 4: equal variance
#### Produce a Residuals-versus-Fitted plot (RvF plot)
```{r}
plot(RESID ~ PRED, data = GLOW_4); abline(h=0)
```

*Conclusion.* No strong evidence of heteroskedasticity.

# Use the model to make predictions
## What is the predicted BMI for women of the following ages: 25, 33, 40, 53, 67, 83 and 98?
### Produce a data frame with *the same variable names* as the model. Include the values of the independent variable.
```{r}
NEWAGE <- c(25, 33, 40, 53, 67, 83, 98)
NEWGLOW <- data.frame(AGE = NEWAGE)
NEWGLOW$NEWBMI <- predict.lm(GLOW_4.LM1, NEWGLOW)
NEWGLOW
```

# Fit a simple linear regression of BMI on SMOKE.F.
```{r}
GLOW_4.LM2 <- lm(BMI ~ SMOKE.F, data = GLOW_4)
summary(GLOW_4.LM2)
```

How do you interpret the coefficient?

# Fit a simple linear regression of BMI on RATERISK.F.
```{r}
GLOW_4.LM3 <- lm(BMI ~ RATERISK.F, data = GLOW_4)
summary(GLOW_4.LM3)
```

Let's vary the reference level of the factor
```{r}
GLOW_4$RATERISK.F <- relevel(GLOW_4$RATERISK.F, ref = 2)
GLOW_4.LM4 <- lm(BMI ~ RATERISK.F, data = GLOW_4)
summary(GLOW_4.LM4)
```

Let's compare coefficients.
```{r}
coefficients(GLOW_4.LM3)
coefficients(GLOW_4.LM4)
```

# THE END