---
title: "Week 12"
author: "Elmer V Villanueva"
date: "May 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Management
## Set Working Directory
```{r}
#setwd("G:/Dropbox/00 - Working Folder/Teaching/DPH112/2017-2018/Week 12")
```

## Install and load the packages `MASS, car, GGally`
```{r}
if (!require("MASS")) install.packages("MASS", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("MASS")

if (!require("car")) install.packages("car", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("car")

if (!require("GGally")) install.packages("GGally", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("GGally")
```


## Read the HINTS data from the `HINTS.csv` file, importing only the following variables: PersonID, WeeklyMinutesModerateExercise, Age, SelfGender, GeneralHealth, MedConditions_HighBP
```{r}
HINTS_5 <- read.csv("HINTS.csv")[ ,c('personid', 'weeklyminutesmoderateexercise', 'age', 'selfgender', 'generalhealth', 'medconditions_highbp')]
str(HINTS_5)
head(HINTS_5)
```

## Recode inappropriate variables
```{r}
HINTS_5$weeklyminutesmoderateexercise <- recode(HINTS_5$weeklyminutesmoderateexercise, "-9=NA")
HINTS_5$age <- recode(HINTS_5$age, "-9=NA; -4=NA")
HINTS_5$selfgender <- recode(HINTS_5$selfgender, "-9=NA; -5=NA")
HINTS_5$generalhealth <- recode(HINTS_5$generalhealth, "-9=NA; -5=NA")
HINTS_5$medconditions_highbp <- recode(HINTS_5$medconditions_highbp, "-9=NA; -1=NA")
str(HINTS_5)
```

## Recode factor variables
```{r}
HINTS_5$selfgender <- factor(HINTS_5$selfgender, labels = c("Male", "Female"))
HINTS_5$generalhealth <- factor(HINTS_5$generalhealth, labels = c("Excellent", "Very Good", "Good", "Fair", "Poor")) 
HINTS_5$generalhealth <- relevel(HINTS_5$generalhealth, ref = 5)
HINTS_5$medconditions_highbp <- factor(HINTS_5$medconditions_highbp, labels = c("Yes", "No"))
str(HINTS_5)
```

# Explore the determinants of weeklyminutesmoderateexercise
## Produce a scatterplot
```{r}
scatterplotMatrix(~ weeklyminutesmoderateexercise + age + selfgender + generalhealth + medconditions_highbp, data = HINTS_5)
ggpairs(columns = 2:6, data = HINTS_5, upper = list(continuous = "points"))
```

### Note the density plot of the dependent variable
```{r}
densityPlot(HINTS_5$weeklyminutesmoderateexercise)
```

## ERROR! Fit a linear regression of weeklyminutesmoderateexercise on the covariates. ERROR!
```{r}
summary(HINTS.WRONG <- lm(weeklyminutesmoderateexercise ~ age + selfgender + generalhealth + medconditions_highbp, data = HINTS_5))
```

## Fit a Poisson regression of weeklyminutesmoderateexercise on the covariates.
```{r}
summary(HINTS.POISS <- glm(weeklyminutesmoderateexercise ~ age + selfgender + generalhealth + medconditions_highbp, data = HINTS_5, family = "poisson"))
```

## Check assumptions and influential points
### Goodness of fit is examined using the deviance.
### Linearity
```{r}
crPlots(HINTS.POISS)
```

### Influential points
```{r}
influence.measures(HINTS.POISS)
influencePlot(HINTS.POISS)
```

### Produce estimates of effect and 95% confidence intervals
```{r}
est <- cbind(Estimate = coef(HINTS.POISS), confint.default(HINTS.POISS))
est
exp(est)
```

# THE END!!


