---
title: "Week 09"
author: "Elmer V Villanueva"
date: "April 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Management

## Install and load the package `car`
```{r}
if (!require("car")) install.packages("car", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("car")
```

## Read the GLOW500 data from the `GLOW500.csv` file, importing only the following variables: SUB_ID, AGE, WEIGHT, HEIGHT, BMI, FRACSCORE 
```{r}
GLOW_5 <- read.csv("GLOW500.csv")[ ,c('SUB_ID', 'AGE', 'WEIGHT', 'HEIGHT', 'BMI', 'FRACSCORE')]
str(GLOW_5)
head(GLOW_5)
```

# Produce a scatterplot matrix
```{r}
scatterplotMatrix(~FRACSCORE + AGE + WEIGHT + HEIGHT + BMI,
                   data = GLOW_5,
                   diag = "boxplot")
```

*Conclusion.* There is no evidence of non-normality, heterogeneity of variance or non-linearity.

# Fit a multiple linear regression of FRACSCORE on AGE, WEIGHT, HEIGHT and BMI.
## Specify the linear model.
```{r}
GLOW_5.LM1 <- lm(FRACSCORE ~ AGE + WEIGHT + HEIGHT + BMI,
                 data = GLOW_5)
```

## Examine the parameter estimates
```{r}
summary(GLOW_5.LM1)
```

*Conclusion.* The model is highly statistically significant (p<0.001). The four variables jointly explain about 76% of the total variability in FRACSCORE. For every year AGE increases, FRACSCORE increases by 0.25 units after controlling for WEIGHT, HEIGHT and BMI. For every 1 unit increase in BMI, FRACSCORE decreases by 0.25 units after adjusting for AGE, WEIGHT and HEIGHT.

## Produce confidence intervals for the parameter estimates
```{r}
confint(GLOW_5.LM1)
```

*Conclusion.* For every year AGE increases, FRACSCORE increases by 0.25 (95% CI 0.23, 0.26) units after controlling for WEIGHT, HEIGHT and BMI.

# Test the assumptions of the model
## Produce the predicted values and the residuals
```{r}
GLOW_5$PRED <- predict(GLOW_5.LM1)
GLOW_5$RESID <- residuals(GLOW_5.LM1)
```

## Test assumption 1: linearity of effect
### Explore plots of the relationships between the response variable (FRACSCORE) and each of the predictor variables while holding the other predictor variables constant.
```{r}
avPlots(GLOW_5.LM1, ask = F)
```

## Test assumption 2: independence
This is not tested statistically.

## Test assumption 3: normality of residuals
```{r}
hist(GLOW_5$RESID, main = "Histogram of Residuals")
plot(density(GLOW_5$RESID), main = "Density of Residuals")
qqnorm(GLOW_5$RESID, main = "Quantile-Quantile Plot of Residuals")
shapiro.test(GLOW_5$RESID)
```

## Test assumption 4: equal variance
### Produce a Residuals-versus-Fitted plot (RvF plot)
```{r}
plot(RESID ~ PRED, data = GLOW_5); abline(h=0)
```

## Test assumption 5: multicollinearity
### Produce a scatterplot matrix of the independent variables
```{r}
scatterplotMatrix(~AGE + WEIGHT + HEIGHT + BMI,
                   data = GLOW_5,
                   diag = "boxplot")
```

### Calculate VIFs
```{r}
vif(GLOW_5.LM1)
```

*Conclusion.* Multicollinearity is a big problem.

Let us omit WEIGHT from the model and test again.
```{r}
GLOW_5.LM2 <- lm(FRACSCORE ~ AGE + HEIGHT + BMI,
                 data = GLOW_5)
GLOW_5$PRED <- predict(GLOW_5.LM2)
GLOW_5$RESID <- residuals(GLOW_5.LM2)

avPlots(GLOW_5.LM2, ask = F)
plot(density(GLOW_5$RESID), main = "Density of Residuals")
plot(RESID ~ PRED, data = GLOW_5); abline(h=0)
vif(GLOW_5.LM2)
```

*Conclusion.* The model assumptions are better met by this reduced model.

The new model has the following estimated parameters
```{r}
summary(GLOW_5.LM2)
summary(GLOW_5.LM1) # For comparison
```

# Examine influential observations.
## Look for outliers. Outliers have studentized residuals less than -2 or greater than 2.
```{r}
GLOW_5$RSTUD <- rstudent(GLOW_5.LM2)
hist(GLOW_5$RSTUD, main = "Histogram of Residuals")
plot(rstudent(GLOW_5.LM2), type = "h"); abline(h=c(-2,2), col=2)
OUTLIERS.LM2 <- subset(GLOW_5,
                       RSTUD < -2 | RSTUD > 2)
OUTLIERS.LM2
```

## Look for leverage. High leverage observations have leverage greater than (2k + 2)/n = (2 * 3 + 2) / 500 = 8/500 = 0.016
```{r}
GLOW_5$HATS <- hatvalues(GLOW_5.LM2)
plot(hatvalues(GLOW_5.LM2), type = "h"); abline(h=0.016, col=2)
HATS.LM2 <- subset(GLOW_5, HATS > 0.016)
HATS.LM2
```

## Look for influential observations. High influence observations have Cook's distance > 4/n = 4/500 = 0.008
```{r}
GLOW_5$COOKSD <- cooks.distance(GLOW_5.LM2)
plot(cooks.distance(GLOW_5.LM2), type = "h"); abline(h=0.008, col=2)
COOKSD.LM2 <- subset(GLOW_5, COOKSD > 0.008)
COOKSD.LM2
```

```{r}
influencePlot(GLOW_5.LM2, id.method="noteworthy", main="Influence Plot", sub="Circle size is proportional to Cook's Distance")
```

# The final model is
```{r}
summary(GLOW_5.LM2)
```

What is wrong with this model?

# THE END

```{r}
GLOW_5.LM3 <- lm(FRACSCORE ~ AGE + BMI, data = GLOW_5)
summary(GLOW_5.LM3)
confint(GLOW_5.LM3)
GLOW_5$PRED <- predict(GLOW_5.LM3)
GLOW_5$RESID <- residuals(GLOW_5.LM3)
avPlots(GLOW_5.LM3, ask = F)
hist(GLOW_5$RESID, main = "Histogram of Residuals")
plot(RESID ~ PRED, data = GLOW_5); abline(h=0)
vif(GLOW_5.LM3)
influencePlot(GLOW_5.LM3, id.method="noteworthy", main="Influence Plot", sub="Circle size is proportional to Cook's Distance")
```




