---
title: "Week 11"
author: "Elmer V Villanueva"
date: "April 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Management
## Install and load the packages `car, plyr, ggplot2, LogisticDx`
```{r}
if (!require("car")) install.packages("car", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("car")

if (!require("plyr")) install.packages("plyr", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("plyr")

if (!require("ggplot2")) install.packages("ggplot2", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("ggplot2")

if (!require("LogisticDx")) install.packages("LogisticDx", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("LogisticDx")

```


## Read the GLOW500 data from the `GLOW500.csv` file, importing only the following variables: SUB_ID, FRACTURE, AGE, BMI, RATERISK, PRIORFRAC
```{r}
GLOW_5 <- read.csv("GLOW500.csv")[ ,c('SUB_ID', 'FRACTURE', 'AGE', 'BMI', 'RATERISK', 'PRIORFRAC')]
GLOW_5$FRACTURE.F <- factor(GLOW_5$FRACTURE, labels = c("No", "Yes"))
GLOW_5$RATERISK.F <- factor(GLOW_5$RATERISK, labels = c("Less", "Same", "More"))
GLOW_5$RATERISK.F <- relevel(GLOW_5$RATERISK.F, ref = 2)
GLOW_5$PRIORFRAC.F <- factor(GLOW_5$PRIORFRAC, labels = c("No", "Yes"))
str(GLOW_5)
head(GLOW_5)
```

# Examine the determinants of FRACTURE.
## Start with AGE
### Plot FRACTURE by AGE
```{r}
plot(FRACTURE ~ AGE, data = GLOW_5)
```

### Replot FRACTURE as a factor variable by AGE
```{r}
plot(FRACTURE.F ~ AGE, data = GLOW_5)
```

### Produce AGE categories
```{r}
AGECAT <- cut(GLOW_5$AGE, 
              breaks = seq(55, 90, by = 5))

FRACAGE <-ddply(GLOW_5,
                .(AGECAT),
                summarise,
                PROP=sum(FRACTURE)/length(FRACTURE))

ggplot(subset(FRACAGE, !is.na(AGECAT)),
       aes(as.factor(AGECAT),
           y=PROP)) +
    geom_bar(stat="identity")
```

### Let's try to fit a LINEAR REGRESSION of FRACTURE.F on AGE
```{r}
FRACAGE.LM1 <- lm(FRACTURE.F ~ AGE, data = GLOW_5)
```

### Let's retry fitting FRACTURE on AGE
```{r}
FRACAGE.LM2 <- lm(FRACTURE ~ AGE, data = GLOW_5)
summary(FRACAGE.LM2)
plot(FRACTURE ~ AGE, data = GLOW_5)
abline(FRACAGE.LM2)
```

# WARNING! This is **TERRIBLY ERRONEOUS**! WARNING!

## Fit a proper logistic model of FRACTURE on AGE.
```{r}
FRACAGE.LM3 <- glm(FRACTURE ~ AGE, 
                   data = GLOW_5,
                   family = binomial)
summary(FRACAGE.LM3)
```

### Test linearity assumption
```{r}
crPlots(FRACAGE.LM3)
```

### Test goodness of fit
```{r}
gof(FRACAGE.LM3)
```

### Examine influential points
```{r}
influence.measures(FRACAGE.LM3)
influencePlot(FRACAGE.LM3)
```

### Produce odds ratios
```{r}
exp(cbind(OR = coef(FRACAGE.LM3), confint.default(FRACAGE.LM3)))
```

## Fit FRACTURE on AGE, BMI, PRIORFRAC, RATERISK
```{r}
FRAC.FULL <- glm(FRACTURE ~ AGE + BMI + PRIORFRAC.F + RATERISK.F, 
                 data = GLOW_5,
                 family = binomial)
summary(glm(FRACTURE ~ AGE + BMI + PRIORFRAC.F + RATERISK.F, 
            data = GLOW_5,
            family = binomial))
summary(glm(FRACTURE ~ AGE + PRIORFRAC.F + RATERISK.F, 
            data = GLOW_5,
            family = binomial))
FRAC.FINAL <- glm(FRACTURE ~ AGE + PRIORFRAC.F + RATERISK.F, 
                  data = GLOW_5, 
                  family = binomial)
```

### Test linearity, goodness of fit and influencial points
```{r}
crPlots(FRAC.FINAL)
gof(FRAC.FINAL)
influence.measures(FRAC.FINAL)
influencePlot(FRAC.FINAL)
```

### Produce odds ratios and confidence intervals.
```{r}
summary(FRAC.FINAL)
exp(cbind(OR = coef(FRAC.FINAL), confint.default(FRAC.FINAL)))
```

# THE END








