---
title: "Week 06"
author: "Elmer V Villanueva"
date: "April 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data management
## Set Working Directory
```{r}
setwd("G:/Dropbox/00 - Working Folder/Teaching/DPH112/2017-2018/Week 06")
```

## Read the GLOW500 data from the `GLOW500.csv` file
```{r}
GLOW_ALL <- read.csv("GLOW500.csv")
str(GLOW_ALL)
head(GLOW_ALL)
```

## Read only the following three variables: AGE, HEIGHT, WEIGHT
```{r}
GLOW_3 <- read.csv("GLOW500.csv")[ ,c('AGE', 'HEIGHT', 'WEIGHT')]
str(GLOW_3)
head(GLOW_3)
```

## Transform HEIGHT to meters
```{r}
GLOW_3$HEIGHTCM <- GLOW_3$HEIGHT/100
str(GLOW_3)
```

## Calculate BMI
```{r}
GLOW_3$BMI <- GLOW_3$WEIGHT/GLOW_3$HEIGHTCM^2
str(GLOW_3)
```

## Drop the HEIGHT variable
```{r}
GLOW_3$HEIGHT <- NULL
str(GLOW_3)
```

## Produce summary statistics
```{r}
summary(GLOW_3)
```

# Basic correlation
```{r}
plot(WEIGHT ~ BMI, data = GLOW_3)
smoothScatter(GLOW_3$BMI, GLOW_3$WEIGHT)
cor(GLOW_3$BMI, GLOW_3$WEIGHT)
cov(GLOW_3$BMI, GLOW_3$WEIGHT)/(sd(GLOW_3$BMI)*sd(GLOW_3$WEIGHT))
```

# Fancy correlation
## Install and load the package `car`
```{r}
if (!require("car")) install.packages("car", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("car")
```

## Proceed with the correlation
```{r}
scatterplot(WEIGHT ~ BMI, data = GLOW_3)
cor.test(~ WEIGHT + BMI, data = GLOW_3)
```

# A word of warning
## Load the Anscombe dataset
```{r}
library("datasets")
str(anscombe)
head(anscombe)
summary(anscombe)
```

## Calculate the correlations of the `x, y` pairs in the Anscombe data
```{r}
cor(anscombe$x1, anscombe$y1)
cor(anscombe$x2, anscombe$y2)
cor(anscombe$x3, anscombe$y3)
cor(anscombe$x4, anscombe$y4)
```

What do you notice?

# THE END