---
title: "Week 10 R Lecture"
author: "Elmer V Villanueva"
date: "11/11/2019"
output: html_document
---

---
# Comment
# setwd("D:/Dropbox/00 - Working Folder/Teaching/DPH101/2019-2020/Week 10 Testing a Single Parameter/R10 R Workshop")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# New functions
- `t.test()`

# Set your working directory
```{r}
setwd("D:/git/DPH101-xjtlu/Y3/week09_lec_11.11")
getwd()
```

# One-sample t-test
Enter the tapeworm data
```{r}
TAPE.1 <- c(14, 9, 18, 26, 12, 0, 10, 4, 8, 21, 28, 24, 24, 2, 3, 14, 9)
```

Visualise your data
```{r}
hist(TAPE.1,
     xlab = "Lung Transit Time (Hours)",
     main = "Duration of Skin Penetration to Lung Appearance \n of Hymenolepis muris")
```

Summarise your data
```{r}
summary(TAPE.1)
```

Perform a one-sample t-test. Assume $\mu = 15$ and $\alpha = 0.01$.
```{r}
t.test(TAPE.1, 
       mu = 15,
       alternative = c("two.sided"),
       conf.level = 0.99)
```

# Independent samples t-test with equal variance
Enter, visualise and summarise the second set of data
```{r}
TAPE.2 <- c(19, 13, 22, 31, 17, 5, 25, 32, 29, 29, 7, 9)
hist(TAPE.2,
     xlab = "Lung Transit Time (Hours)",
     main = "Duration of Skin Penetration to Lung Appearance \n of the Second Sample of Hymenolepis muris")
summary(TAPE.2)
```

Perform an independent samples t-test with equal variance. Assume $\alpha = 0.05$.
```{r}
t.test(TAPE.1, TAPE.2,
       alternative = c("two.sided"),
       var.equal = TRUE,
       conf.level = 0.95)
```

# Independent samples t-test with unequal variance
Enter, visualise and summarise the third set of data
```{r}
TAPE.3 <- c(91, 13, 22, 31, 17, 5, 25, 32, 29, 29, 7, 9)
hist(TAPE.3,
     xlab = "Lung Transit Time (Hours)",
     main = "Duration of Skin Penetration to Lung Appearance \n of the Third Sample of Hymenolepis muris")
summary(TAPE.3)
```

Perform an independent samples t-test with unequal variance. Assume $\alpha = 0.05$.
```{r}
t.test(TAPE.1, TAPE.3,
       alternative = c("two.sided"),
       var.equal = FALSE,
       conf.level = 0.95)
```

# Paired samples t-test 
Enter, visualise and summarise the data
```{r}
TAPE.BEFORE <- c(20.0, 63.3, 96.0, 9.2, 3.1, 50.0, 33.0, 69.0)
TAPE.AFTER <- c(63.5, 91.5, 59.0, 37.8, 10.1, 19.6, 41.0, 87.8)
TAPE.DELTA <- TAPE.AFTER - TAPE.BEFORE

hist(TAPE.DELTA,
     xlab = "Lung Function",
     main = "Difference in Lung Function Following  \n Infection with Hymenolepis muris")
summary(TAPE.DELTA)
```

Perform a paired samples t-test. Assume $\alpha = 0.10$.
```{r}
t.test(TAPE.AFTER, TAPE.BEFORE,
       alternative = c("two.sided"),
       paired = TRUE,
       conf.level = 0.90)
```

# THE END