---
title: "DPH101/ENV203 Assignment Basic Probability ANSWERS"
author: "Elmer V Villanueva"
date: "October 07, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

Given

##### Table 1. Skill retention by intervention group.
| Competence | CPR | CC-CPR | Total |
|:-----------|----:|-------:|------:|
| Yes        | 12  | 15     | 27    |
| No         | 15  | 14     | 29    |
| Total      | 27  | 29     | 56    |

Find the following probabilities:

1. $P(CC-CPR) = 29/56 = 0.5178$
```{r}
29/56
```

2. $P(Competent) = 27/56 = 0.4821$
```{r}
27/56
```

3. $P(Competent \cap CPR) = P(Competent | CPR) \cdot P(CPR) = (12/27) \cdot (27/56) = 12/56 = 0.2143$
```{r}
12/56
```

4. $P(Competent \cup CC-CPR) = P(Competent) + P(CC-CPR) - P(Competent \cap CC-CPR) = 27/56 + 29/56 - 15/56 = 41/56 = 0.7321$
```{r}
41/56
```

5. $P(Competent | CC-CPR) = P(Competent \cap CC-CPR) / P(CC-CPR) = (15/56) / (29/56) = 15/29 = 0.5172$
```{r}
15/29
```

# Question 2

$P(age > 30 \cup master's\ degree) = P(age > 30) + P(master's\ degree) - P(age > 30 \cap master's\ degree) = 15/25 + 8/25 - 2/25 = 21/25 = 0.8400$
```{r}
21/25
```

# Question 3
The definition of a conditional probability is given as 

$$ P(B|A) = \dfrac{P(B \cap A)}{P(A)};\ P(A) \neq 0$$

If we specify events B and A to be {allergy} and {exposed}, respectively, and solve for $P(exposed)$, we get:

$$
P(exposed) = \dfrac{P(allergy \cap exposed)}{P(allergy | exposed)} = \dfrac{0.6}{0.8} = 0.7500
$$
```{r}
0.6/0.8
```

# THE END