---
title: "Week 05 R Workshop"
author: "Elmer V Villanueva"
date: "October 7, 2019"
output:
  pdf_document: default
  html_document: default
---

---
# Comment
# setwd("D:/Dropbox/00 - Working Folder/Teaching/DPH101/2019-2020/Week 05 Discrete Probability Distributions/R05 R Workshop")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# New functions and packages
- `choose()`  c(n x)    combination
- `sum()`
- `dbinom()`  calculate P(probability) of x
- `pbinom()`  calculate F(cumulative probability) of x
- `dpois()`  calculate P(probability) of x
- `ppois()`  calculate F(cumulative probability) of x

# Set your working directory
```{r}
setwd("D:/git/DPH101-xjtlu/Y3/week05_lec_10.7")
```

# Load GLOW500 Data
```{r}
GLOW500_WORK <- read.csv("GLOW500.csv")
str(GLOW500_WORK)
head(GLOW500_WORK)
```

Let's focus on the variable `PRIORFRAC`. It's still a numeric variable and we need to transform it into a factor variable.
```{r}
GLOW500_WORK$PRIORFRAC.F <- factor(GLOW500_WORK$PRIORFRAC, 
                                  labels=c("No", "Yes"))
table(GLOW500_WORK$PRIORFRAC.F, GLOW500_WORK$PRIORFRAC)
```

## Question 1: What is the probability of a history of prior fracture in this study population?
```{r}
prop.table(table(GLOW500_WORK$PRIORFRAC.F))
```

-----

# Binomial Distribution

## Question 2: Say we select 10 women at random from the study population. What is the probability that exactly four of them have a history of prior fracture?

Prior fracture is a binomial process. First, the probability of prior fracture is the same for every woman. Second, the prior fracture history of one woman is not related to any other women. 

The history of prior fracture is given as $126 / 500 = 0.2520$. In addition, we have selected 10 women. These two pieces of information give use out parameters, $p$ and $n$, respectively.

Let us visualise this problem.

```{r echo=FALSE}
SUCCESS.B <- c(0:10)
PROBABILITY.B <- dbinom(0:10, 10, 0.2520)
barplot(height = PROBABILITY.B,
        names = SUCCESS.B,
        col = c(rep("grey", times = 4), "red", rep("grey", times = 7)),
        ylim = c(0, 0.3),
        ylab = "Probability",
        xlab = "Number with Prior Fracture History",
        main = "Probability Distribution (Bin(10, 0.2520))")
```

Finally, to calculate the probability, we can use the probability expression $P(X = x) = C\binom{n}{x}p^x(1-p)^{(n-x)}$.
```{r}
P <- 126/500; N <- 10; X <- 4
choose(N, X) * P ^ X * (1-P) ^ (N-X)
```

Another way to calculate this is to use the `dbinom` function, which gives the *density* or $P(X = x)$.
```{r}
dbinom(4, 10, 0.2520) # success, all trial, probability
```

## Question 3: What is the probability that exactly five out of 10 have a prior fracture history?

Let's visualise the problem.

```{r echo=FALSE}
barplot(height = PROBABILITY.B,
        names = SUCCESS.B,
        col = c(rep("grey", times = 5), "red", rep("grey", times = 6)),
        ylim = c(0, 0.3),
        ylab = "Probability",
        xlab = "Number with Prior Fracture History",
        main = "Probability Distribution (Bin(10, 0.2520))")
```

To solve, 
```{r}
dbinom(5, 10, 0.2520)
```

## Question 4: What is the probability that fewer than three out of 10 have a prior history of fracture?
```{r echo=FALSE}
barplot(height = PROBABILITY.B,
        names = SUCCESS.B,
        col = c(rep("red", times = 3), rep("grey", times = 8)),
        ylim = c(0, 0.3),
        ylab = "Probability",
        xlab = "Number with Prior Fracture History",
        main = "Probability Distribution (Bin(10, 0.2520))")
```

We can solve this problem a number of ways. First, we can summate using brute force:
```{r}
dbinom(0, 10, 0.2520) + dbinom(1, 10, 0.2520) + dbinom(2, 10, 0.2520)
```

Second, we can use the `sum` function with a sequence operator:
```{r}
sum(dbinom(0:2, 10, 0.2520))
```

Third, we can use the `pbinom' function, which gives the *distribution function* or $P(X \le x)$.
```{r}
pbinom(2, 10, 0.2520)
```

## Question 5: What is the probability that two out of fourteen randomly selected women have a prior history of fracture?

-----

# Poisson Distribution

## Question 6: What is the probability that exactly four randomly selected women have a prior fracture?

In this question, we can assume that the events are distributed as a Poisson process and with parameter $\lambda = 0.2520$. 

Let us visualise this problem.

```{r echo=FALSE}
SUCCESS.P <- c(0:10)
PROBABILITY.P <- dpois(0:10, 0.2520)
barplot(height = PROBABILITY.P,
        names = SUCCESS.P,
        col = c(rep("grey", times = 4), "red", rep("grey", times = 7)),
        ylim = c(0, 0.8),
        ylab = "Probability",
        xlab = "Number with Prior Fracture History",
        main = "Probability Distribution (Pois(0.2520))")
```

The probability is calculated as $P(X=x) = e^{-\lambda} \lambda^x / x!$. In this problem, $x = 4$.
```{r}
X <- 4; LAMBDA <- 0.2520
exp(-LAMBDA) * LAMBDA ^ X / factorial(X)
```

Similar to the binomial density function, `dpois` provides a Poisson density function:
```{r}
dpois(4, 0.2520)
```

## Question 7: What is the probability that fewer than three randomly selected women have a prior fracture?
```{r echo=FALSE}
barplot(height = PROBABILITY.P,
        names = SUCCESS.P,
        col = c(rep("red", times = 3), rep("grey", times = 8)),
        ylim = c(0, 0.8),
        ylab = "Probability",
        xlab = "Number with Prior Fracture History",
        main = "Probability Distribution (Pois(0.2520))")
```

```{r}
dpois(0, 0.2520) + dpois(1, 0.2520) + dpois(2, 0.2520)
sum(dpois(0:2, 0.2520))
ppois(2, 0.2520)
```

## Question 8: What is the probability that fourteen randomly selected women have a prior history of fracture?

# THE END






