---
title: "Week 03 Assignment"
author: "1717719"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let us produce numerical summaries of the variables AGE, HEIGHT and WEIGHT in the GLOW500 data.
First, make sure to set your working directory.

```{r}
setwd("D:/git/DPH101-xjtlu/Y3/week03_ass_9.23")
```

Then, read the file into memory.
```{r}
GLOW500_WORK <- read.csv("GLOW500.csv")
```


We will calculate the mean and variance for the three variables. The formula for the mean is $\bar{x} = \frac{\sum{x}}{n}$. 

The formula for the variance is $\sigma^{2} = \frac{\sum{(x - \bar x)^2}}{n-1}$.

The mean and variance of the three variables are calculated using base R functions. Let's demonstrate using AGE.

```{r}
mean(GLOW500_WORK$AGE); var(GLOW500_WORK$AGE)
```


\begin{table}[h!]
\caption{Basic numerical summaries of age, weight and height in the GLOW500 study.}
\centering
\begin{tabular}{lrr}
\hline
Variable  & Mean        & Variance \\
\hline
Age(Year)      & 68.6      & 80.81      \\
Weight(kg)     & 71.82     & 270.142      \\
Height(cm)     & 161.4     & 40.39      \\
\hline
\end{tabular}
\end{table}


Finally, let's visualise the relationship among all three variables using a scatterplot matrix.

```{r}
pairs(~AGE + WEIGHT + HEIGHT, data=GLOW500_WORK, main="Simple Scatterplot Matrix")
```


# THE END

