---
title: "Week 03 Assignment"
author: "< Insert Student ID >"
date: "23 September 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let us produce numerical summaries of the variables `AGE`, `HEIGHT` and `WEIGHT` in the `GLOW500` data.

First, make sure to set your working directory.
```{r}
setwd("D:/git/DPH101-xjtlu/Y3/week03_ass_9.23")
```

Then, read the file into memory.
```{r echo=FALSE}
GLOW500_WORK <- read.csv("GLOW500.csv")
```

We will calculate the mean and variance for the three variables. The formula for the mean is $\bar{x} = \frac{\Sigma x}{n}$. The formula for the variance is $\sigma^2 = \frac{\Sigma{(x - \bar{x})^2}}{n-1}$.

The mean and variance of the three variables are calculated using base R functions. Let's demonstrate using `AGE`.
```{r}
mean(GLOW500_WORK$AGE); var(GLOW500_WORK$AGE)
```

We use similar commands for `WEIGHT` and `HEIGHT`.
```{r include=FALSE}
mean(GLOW500_WORK$WEIGHT); var(GLOW500_WORK$WEIGHT)
mean(GLOW500_WORK$HEIGHT); var(GLOW500_WORK$HEIGHT)
```

Then, we can produce a table to summarise the results.

##### Table 1. Basic numerical summaries of age, weight and height in the GLOW500 study. 
| Variable    |  Mean  | Variance |
|:------------|-------:|---------:|
| Age (years) | 68.6   | 80.81    |
| Weight (kg) | 71.82  | 270.142  |
| Height (cm) | 161.4  | 40.39    |

Finally, let's visualise the relationship among all three variables using a scatterplot matrix.
```{r}
pairs(~AGE + WEIGHT + HEIGHT, data=GLOW500_WORK,
      main="Simple Scatterplot Matrix")
```

# THE END