---
title: '1717719'
author: "dizhen"
date: "2019/9/15"
output:
  html_document: default
  pdf_document: default
---
Set working directory and import the dataset:
```{r}
setwd("D:/git/DPH101-xjtlu/Y3/week02_ass_9.16")
IQLEAD_table <- read.table("IQLEAD.txt", header = TRUE)
save(IQLEAD_table, file="IQLEAD.RData")
```

# Question 1: 

Prepare a properly formatted table
```{r}
View(IQLEAD_table)
head(IQLEAD_table)
```
Summarises the ages of the subjects.
```{r}
summary(IQLEAD_table$AGE)
```
The table's stub must use classes of width three.
```{r}
IQLEAD_table$AGECAT <- cut(IQLEAD_table$AGE, breaks = seq(2, 17, by = 3))
summary(IQLEAD_table$AGECAT)
```
The second column must present the frequencies, The third column must present the relative frequencies, The last column must present the cumulative relative frequencies.
```{r}
if (!require("summarytools")) install.packages("summarytools", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("summarytools")
freq(IQLEAD_table$AGECAT)
```

```{r}
tr_table <- freq(IQLEAD_table$AGECAT)[,c(1:3)]
tr_table <- as.data.frame(tr_table)
write.csv(tr_table,"freq_table.csv")
```

# Question 2:

Construct a properly formatted histogram of the ages of the subjects using bin widths of three.
```{r}
png(filename="histogram of age.png")
hist(IQLEAD_table$AGE,
     xlab = "Age (years)",
     main = "Histogram of Age in IQLEAD",
     border = "white",
     col = "dark blue",
     breaks = seq(2, 17, by = 3))
dev.off()
```

# Question 3:

Construct a properly formatted strip plot of the ages of the subjects.
```{r}
png(filename = "stripchart of age.png")
stripchart(IQLEAD_table$AGE, 
           method = "stack", 
           at = 0.3,
           pch = 16,
           xlab = "Age (years)",
           main = "Distribution of Age in IQLEAD")
dev.off()
```

# Question 4:
Construct a properly formatted Pareto plot of the sexes of the subjects.
```{r}

IQLEAD_table$F_SEX <- factor(IQLEAD_table$SEX, 
                                 labels=c("Male", "Female"))
table(IQLEAD_table$F_SEX, IQLEAD_table$SEX)

freq_table <- freq(IQLEAD_table$F_SEX)[c(1:2),c(1)]
png(filename = "pareto chart of sex.png")
barplot(freq_table, ylab = "Frequency", xlab = "Sex")
dev.off()
```

