---
title: "Week 03 R Workshop"
author: "Elmer V Villanueva"
date: "September 16, 2019"
output: pdf_document
---

---
# Comment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SET YOUR WORKING DIRECTORY!
```{r}
setwd("D:/git/DPH101-xjtlu/Y3/week03_lec_9.16")
```

# Load the `GLOW500` data.
```{r}
GLOW500_WORK <- read.csv("GLOW500.csv")
```

Make sure that you develop the habit of checking that the file was loaded correctly.
```{r}
str(GLOW500_WORK)
head(GLOW500_WORK)
```

# Boxplots
Let's produce a boxplot of `AGE`.
```{r}
boxplot(GLOW500_WORK$AGE,
        ylab = "Age (years)",
        main = "Distribution of Age in GLOW500")
```

Let's produce a boxplot of `WEIGHT` by prior fracture status (`PRIORFRAC`).
```{r}
boxplot(WEIGHT ~ PRIORFRAC, data = GLOW500_WORK,
        ylab = "Weight (kg)",
        xlab = "Prior Fracture",
        main = "Distribution of Weight by Prior Fracture in GLOW500")
```

Why is the prior fracture labels appearing as zero and one?

# Scatterplot
Let's look at the relationship between `WEIGHT` and `HEIGHT`
```{r}
plot(WEIGHT ~ HEIGHT, data = GLOW500_WORK,
     ylab = "Weight (kg)",
     xlab = "Height (cm)",
     main = "Weight versus Height in GLOW500")
```

# Simple numerical summaries
Let's produce simple numerical summaries of `HEIGHT`.
```{r}
summary(GLOW500_WORK$HEIGHT)
```

## Geometric mean
R doesn't have a built-in function for the geometric mean. However, there is a function in the `EnvStats` package that we can use.
```{r}
if (!require("EnvStats")) install.packages("EnvStats", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("EnvStats")
```

Let's calculate the geometric mean of `WEIGHT`.
```{r}
geoMean(GLOW500_WORK$WEIGHT)
```

There is an easier way, however, that we can do this without installing and loading a package.
```{r}
exp(mean(log(GLOW500_WORK$WEIGHT)))
```

## Harmonic mean
R doesn't have a built-in function to calculate the harmonic mean, but there is a function in the `lmomco` package that we can use.
```{r}
if (!require("lmomco")) install.packages("lmomco", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("lmomco")
```

Let's calculate the harmonic mean of `HEIGHT`.
```{r}
harmonic.mean(GLOW500_WORK$HEIGHT)
```

As before, there is an easier way to calcualte this without needing to load a package.
```{r}
1/mean(1/GLOW500_WORK$HEIGHT)
```

## Quantiles
Let's calcualte the 3rd, 45th, 59th and 93rd percentile of `AGE`.
```{r}
quantile(GLOW500_WORK$AGE, 
         probs = c(0.03, 0.45, 0.59, 0.93))
```

Let's calculate the interquartile range of `AGE`.
```{r}
quantile(GLOW500_WORK$AGE, 
         probs = c(0.25, 0.75))
```

## Standard deviation
Calculate the standard deviation of `HEIGHT`.
```{r}
sd(GLOW500_WORK$HEIGHT)
```

## Variance
Calculate the variance of `WEIGHT`.
```{r}
var(GLOW500_WORK$WEIGHT)
```

## Coefficient of variation
Calculate the coefficeint of variation of `HEIGHT`.
```{r}
sd(GLOW500_WORK$HEIGHT)/mean(GLOW500_WORK$HEIGHT)
```

# Manipulating data
Let's create a small data frame.
```{r}
ID <- c(1, 2, 3, 4, 5)
PETALS <- c(30, 35, 26, 23, 41)
COLOR <- c("Red", "White", "White", "Red", "Red")
ROSE <- data.frame(ID, PETALS, COLOR)
str(ROSE)
head(ROSE)
```

Let's arrange the data in ascending order of the number of petals.
```{r}
ROSE[order(ROSE$PETALS),]
```

Let's produce a dataset containing only data from the red roses.
```{r}
ROSE.RED <- subset(ROSE, COLOR=="Red")
head(ROSE.RED)
```

# THE END