---
title: "Week 12 R Lecture"
author: "Elmer V Villanueva"
date: "11/25/2019"
output: html_document
---

---
# Comment
# setwd("D:/Dropbox/00 - Working Folder/Teaching/DPH101/2019-2020/Week 12 Analysis of Variance/L12 Analysis of Variance")
# info: http://www.sthda.com/english/wiki/two-way-anova-test-in-r
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("D:/git/DPH101-xjtlu/Y3/week12_lec_11.25")
```

# Oneway ANOVA
Let's enter the data.
```{r}
HAEM.CONTROL <- c(38, 40, 32, 36, 40, 40, 38, 40, 38, 40, 36, 40, 40, 35, 45)
HAEM.DUSHU <- c(56, 60, 50, 50, 50, 35, 40, 40, 55, 35, rep(c(NA), times = 5))
HAEM.JINJI <- c(40, 42, 38, 46, 36, rep(c(NA), times = 10))
HAEM.DATA.1 <- data.frame(HAEM.CONTROL, HAEM.DUSHU, HAEM.JINJI)
str(HAEM.DATA.1)
head(HAEM.DATA.1)
View(HAEM.DATA.1)
```

Let's enter the data in another way.
```{r}
HAEMATOCRIT <- c(38, 40, 32, 36, 40, 40, 38, 40, 38, 40, 36, 40, 40, 35, 45,
                 56, 60, 50, 50, 50, 35, 40, 40, 55, 35,
                 40, 42, 38, 46, 36)
LOCATION <- factor(c(rep(0, times = 15),
                     rep(1, times = 10),
                     rep(2, times = 5)),
                   labels = c("Control", "Dushu", "Jinji"))
HAEM.DATA.2 <- data.frame(HAEMATOCRIT, LOCATION)
str(HAEM.DATA.2)
head(HAEM.DATA.2)
View(HAEM.DATA.2)
```

Draw pictures. This graph is incomplete.
```{r}
boxplot(HAEMATOCRIT ~ LOCATION, 
        data = HAEM.DATA.2)
```

Perform the analysis.
```{r}
HAEMATOCRIT.AOV <- aov(HAEMATOCRIT ~ LOCATION, 
                       data = HAEM.DATA.2)
summary(HAEMATOCRIT.AOV)
```

Where is the difference?
```{r}
with(data = HAEM.DATA.2,
     t.test(HAEMATOCRIT[LOCATION == "Control"], HAEMATOCRIT[LOCATION == "Dushu"]))
with(data = HAEM.DATA.2,
     t.test(HAEMATOCRIT[LOCATION == "Control"], HAEMATOCRIT[LOCATION == "Jinji"]))
with(data = HAEM.DATA.2,
     t.test(HAEMATOCRIT[LOCATION == "Dushu"], HAEMATOCRIT[LOCATION == "Jinji"]))
```

Control for multiple looks.
```{r}
TukeyHSD(x=HAEMATOCRIT.AOV)
```

# Twoway ANOVA: Randomised complete block design
Let's enter the data in long form.
```{r}
WEIGHT <- c(70, 80, 90, 100, 110,
            90, 90, 90, 90, 120,
            100, 100, 120, 120, 140)
DRUG <- factor(c(rep(0, times = 5), 
                 rep(1, times = 5), 
                 rep(2, times = 5)),
               labels = c("Control", "Drug A", "Drug B"))
AGE <- factor(c(rep(c(0:4), times = 3)),
              labels = c("<20", "20-29", "30-39", "40-49", "50+"),
              ordered = TRUE)
CHICKEN <- data.frame(WEIGHT, DRUG, AGE)
str(CHICKEN)
head(CHICKEN)
View(CHICKEN)
```

Draw pictures
```{r}
boxplot(WEIGHT ~ DRUG,
        data = CHICKEN)
```

Perform the analysis
```{r}
CHICKEN.AOV <- aov(WEIGHT ~ DRUG + AGE, 
                       data = CHICKEN)
summary(CHICKEN.AOV)
```

```{r}
op <- par(mar = c(4, 4, 0.1, 0.1))
plot(WEIGHT ~ as.numeric(AGE), 
     data = CHICKEN, 
     type = "n", 
     axes = F, 
     xlab = "", 
     ylab = "")
with(subset(CHICKEN, DRUG == "Control"), 
     points(WEIGHT ~ as.numeric(AGE), 
            pch = 21, 
            type = "o", 
            lwd = 1))
with(subset(CHICKEN, DRUG == "Drug A"), 
     points(WEIGHT ~ as.numeric(AGE), 
            pch = 16, 
            type = "o", 
            lwd = 1))
with(subset(CHICKEN, DRUG == "Drug B"), 
     points(WEIGHT ~ as.numeric(AGE), 
            pch = 5, 
            type = "o", 
            lwd = 1, 
            lty = 2))
axis(1, cex.axis = 0.8)
mtext(text = "Age", 
      side = 1, 
      line = 3)
axis(2, cex.axis = 0.8, las = 1)
mtext(text = "Weight", 
      side = 2, 
      line = 3)
legend("topleft", 
       leg = c("Control", "Drug A", "Drug B"), 
       lty = 0, 
       pch = c(21, 16, 5), 
       bty = "n", 
       cex = 0.7)
box(bty = "l")
```

# Twoway ANOVA: Factorial design
Let's enter the data.
```{r}
SIZE <- c(20, 25, 22, 27, 21, 25, 30, 29, 28, 30,
          24, 28, 24, 25, 30, 28, 31, 26, 29, 32,
          30, 35, 30, 35, 36, 30, 29, 31, 30, 30,
          39, 42, 36, 42, 40, 40, 45, 50, 45, 60,
          31, 30, 40, 35, 30, 32, 35, 30, 40, 30,
          41, 46, 40, 40, 35, 42, 50, 40, 55, 45,
          20, 21, 20, 20, 19, 23, 25, 28, 30, 31,
          24, 25, 30, 26, 23, 29, 30, 28, 27, 30)
WHEAT <- factor(c(rep(0, times = 20), rep(1, times=20),
                  rep(2, times = 20), rep(3, times = 20)),
                labels = c("T. aestivum", "T. compactum",
                           "T. dicoccon", "T. urartu"))
d1 <- c(0:3); d2 <- c(rep(d1, each = 5))
DOSE <- factor(c(rep(d2, times = 4)),
               labels = c("Low", "Medium", "High", "Very High"),
               ordered = TRUE)
SELENIUM <- data.frame(SIZE, WHEAT, DOSE)
str(SELENIUM)
head(SELENIUM)
View(SELENIUM)
```

```{r}
boxplot(SIZE ~ WHEAT * DOSE,
        data = SELENIUM)
```

Perform the analysis
```{r}
SELENIUM.AOV <- aov(SIZE ~ WHEAT * DOSE,
                    data = SELENIUM)
summary(SELENIUM.AOV)
```


