---
title: "Week 02"
author: "Elmer V Villanueva"
date: "September 28, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SET YOUR WORKING DIRECTORY!
```{r}
setwd("D:/r project/Y3-biostats/dph101.9.9")
```

# Working with Packages
Load the "ChickWeight" dataset. The dataset is included in R and is used for standard analysis
```{r}
data("ChickWeight")
```

View the help file associated with the dataset
```{r}
help("ChickWeight")
```

If a package does not appear on the list, you need to install it first.
```{r}
if (!require("fortunes")) install.packages("fortunes", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
```

Having installed a package does not mean it has been loaded.
```{r}
#fortunes(256)
```

You need to load it first
```{r}
library(fortunes)
fortune(23)
```

To unload a package, use the `detach()` function
```{r}
detach(package:fortunes)
```

To completely uninstall a package, use the `remove.packages()` function
```{r}
remove.packages("fortunes")
```

***

# Working with Data
Let's enter data into R and produce a data frame.
```{r}
STUDENT <- c(1004, 1855, 2464, 2986)
SEX <- c("M", NA, "F", "M")
GRADE <- c(89, 65, NA, 76)
DPH101 <- data.frame(STUDENT, SEX, GRADE)
```

You can use R's built-in data viewer to inspect and edit data
```{r}
View(DPH101)
```

You can use R's built-in data editor to inspect and edit data
```{r}
edit(DPH101)
```

You can save the data into a directory
```{r}
save(DPH101, file="D:/r project/Y3-biostats/dph101.9.9/DPH101.RData")
```

In order to simplify saving and loading tasks, set the working directory
```{r}
setwd("D:/r project/Y3-biostats/dph101.9.9")
save(DPH101, file="DPH101.RData")
```

We can load the file using the load() function
```{r}
load("DPH101.RData")
```

Now, open the file `GLOW500.csv` in Excel. We need an easier way to enter this into R
```{r}
GLOW500_WORK <- read.csv("GLOW500.csv")
save(GLOW500_WORK, file="GLOW500.RData")
```

Check the data frame
```{r}
View(GLOW500_WORK)
```

Let's look at one of the variables PRIORFRAC and compare it to the codebook
```{r}
GLOW500_WORK$PRIORFRAC
```

Note that this is a numerical variable
```{r}
class(GLOW500_WORK$PRIORFRAC)
```

We want to change this variable type to a factor variable. Before we do this in the actual dataset, let's test this with the variable `TWEAK`.
```{r}
TWEAK <- factor(GLOW500_WORK$PRIORFRAC)
class(TWEAK)
TWEAK
```

We see that `TWEAK` is a factor variable with levels of "0" and "1". These levels are meaningless without the codebook. Let's change that by adding labels.
```{r}
TWEAK <- factor(GLOW500_WORK$PRIORFRAC, 
                labels=c("No", "Yes"))
class(TWEAK)
TWEAK
```

Now that we know it works, we can do this in the real dataset.
```{r}
GLOW500_WORK$PRIORFRAC1 <- factor(GLOW500_WORK$PRIORFRAC, 
                                  labels=c("No", "Yes"))
class(GLOW500_WORK$PRIORFRAC1)
```

Let's check that PRIORFRAC and PRIORFRAC1 contain the same data
```{r}
table(GLOW500_WORK$PRIORFRAC, GLOW500_WORK$PRIORFRAC1)
```

Try revising PREMENO
```{r}
class(GLOW500_WORK$PREMENO)
GLOW500_WORK$PREMENO1 <- factor(GLOW500_WORK$PREMENO, 
                                labels=c("No", "Yes"))
class(GLOW500_WORK$PREMENO1)
table(GLOW500_WORK$PREMENO, GLOW500_WORK$PREMENO1)
```

## Try revising all other factor variables MOMFRAC, ARMASSIST, SMOKE, RATERISK, FRACTURE.
You should try this on your own. Below, however, I have supplied the answers

### Revising MOMFRAC
```{r}
GLOW500_WORK$MOMFRAC1 <- factor(GLOW500_WORK$MOMFRAC, 
                                labels=c("No", "Yes"))
table(GLOW500_WORK$MOMFRAC, GLOW500_WORK$MOMFRAC1)
```

### Revising ARMASSIST
```{r}
GLOW500_WORK$ARMASSIST1 <- factor(GLOW500_WORK$ARMASSIST, 
                                  labels=c("No", "Yes"))
table(GLOW500_WORK$ARMASSIST, GLOW500_WORK$ARMASSIST1)
```

### Revising SMOKE
```{r}
GLOW500_WORK$SMOKE1 <- factor(GLOW500_WORK$SMOKE, 
                              labels=c("No", "Yes"))
table(GLOW500_WORK$SMOKE, GLOW500_WORK$SMOKE1)
```

### Revising FRACTURE
```{r}
GLOW500_WORK$FRACTURE1 <- factor(GLOW500_WORK$FRACTURE, 
                                 labels=c("No", "Yes"))
table(GLOW500_WORK$FRACTURE, GLOW500_WORK$FRACTURE1)
```

### Revising RATERISK
```{r}
GLOW500_WORK$RATERISK1 <- factor(GLOW500_WORK$RATERISK, 
                                 labels=c("Less", "Same", "Greater"))
table(GLOW500_WORK$RATERISK, GLOW500_WORK$RATERISK1)
```

Let's drop some superfluous variables
```{r}
GLOW500_WORK[,"PRIORFRAC"] <- list(NULL)
GLOW500_WORK[,"PREMENO"] <- list(NULL)
GLOW500_WORK[,"MOMFRAC"] <- list(NULL)
GLOW500_WORK[,"ARMASSIST"] <- list(NULL)
GLOW500_WORK[,"SMOKE"] <- list(NULL)
GLOW500_WORK[,"RATERISK"] <- list(NULL)
GLOW500_WORK[,"FRACTURE"] <- list(NULL)
```

Save the dataset as `GLOW500FINAL.RData`.
```{r}
save(GLOW500_WORK, file="GLOW500FINAL.RData")
```

***

# Replicating TTR
## Fequency Distributions
Consider the variable `AGE` in `GLOW500`.
```{r}
summary(GLOW500_WORK$AGE)
```

We can ask R to create age groups as follows:
```{r}
GLOW500_WORK$AGECAT <- cut(GLOW500_WORK$AGE, breaks=c(54,59,64,69,74,79,84,89,Inf))
GLOW500_WORK$AGECAT
```

The labels are awkward. Let's change them.
```{r}
GLOW500_WORK$AGECAT <- cut(GLOW500_WORK$AGE,
                           breaks=c(54,59,64,69,74,79,84,89,Inf),
                           labels=c("55-59", "60-64", "65-69", "70-74", 
                                    "75-79", "80-84", "85-89", "90"))
summary(GLOW500_WORK$AGECAT)
```

The summary gives a frequency distribution, but not a relative frequency distribution. There are two other functions that might help us: `table()` and `prop.table()`. Let's see if it will work:
```{r}
table(GLOW500_WORK$AGECAT)
prop.table(table(GLOW500_WORK$AGECAT))
```

The calculations are there, but they're very hard to understand. 

A neat little function is `freq()` in the `summarytools` package.
```{r}
if (!require("summarytools")) install.packages("summarytools", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library("summarytools")
freq(GLOW500_WORK$AGECAT)
```

## Histograms
Base R's `hist()` function can draw a basic histogram.
```{r}
hist(GLOW500_WORK$AGE)
```

This graph is not properly formatted. Let's change the horizontal axis.
```{r}
hist(GLOW500_WORK$AGE,
     xlab = "Age (years)")
```

Let's include a title.
```{r}
hist(GLOW500_WORK$AGE,
     xlab = "Age (years)",
     main = "Histogram of Age in GLOW500")
```

We can adjust other elements in the graph.
```{r}
hist(GLOW500_WORK$AGE,
     xlab = "Age (years)",
     main = "Histogram of Age in GLOW500",
     border = "red",
     col = "blue")
```

Let's change the size of the breaks.
```{r}
hist(GLOW500_WORK$AGE,
     xlab = "Age (years)",
     main = "Histogram of Age in GLOW500",
     border = "red",
     col = "blue",
     breaks = c(55, 65, 75, 85, 90))
```

# Dot plots
These are better known as strip charts or strip charts.
```{r}
stripchart(GLOW500_WORK$AGE)
```

This doesn't look like much of anything. This is because all the points are overlaid on each other. You need to specify the option `method = "stack"` to produce the effect we want.
```{r}
stripchart(GLOW500_WORK$AGE, 
           method = "stack")
```

The chart is truncated at the top, so we need to lower the chart area.
```{r}
stripchart(GLOW500_WORK$AGE, 
           method = "stack", 
           at = 0.3)
```

We can change the plotting symbol into filled circles.
```{r}
stripchart(GLOW500_WORK$AGE, 
           method = "stack", 
           at = 0.3,
           pch = 16)
```

Finally, let's add an axis label and a title.
```{r}
stripchart(GLOW500_WORK$AGE, 
           method = "stack", 
           at = 0.3,
           pch = 16,
           xlab = "Age (years)",
           main = "Distribution of Age in GLOW500")
```

# Stem plot
This is better known as a stem-and-leaf plot.
```{r}
stem(GLOW500_WORK$AGE)
```

# THE END
